---
output:
  pdf_document: default
  html_document: default
---
# R codes for coursework
## prepare data before solving questions
### set the working directory that contains the files
getwd()
setwd("C:/Users/a0972/Desktop/st2195/coursework")
library(dplyr)
library(DBI)
### create the database
if (file.exists("coursework1.db")) 
  file.remove("courseworl1.db")
conn <- dbConnect(RSQLite::SQLite(), "coursework1.db")

### create tables(read from csv files)
firsty <- read.csv("2007.csv", header = TRUE)
secondy <- read.csv("2008.csv", header = TRUE)
plane <- read.csv("plane-data.csv", header = TRUE)
airport_code<- read.csv("airports.csv", header = TRUE)

dbWriteTable(conn, "firsty", firsty,overwrite = FALSE)
dbWriteTable(conn, "secondy", secondy,overwrite = FALSE)
dbWriteTable(conn, "plane", plane,overwrite = FALSE)
dbWriteTable(conn, "airport_code", airport_code, overwrite = FALSE)

df12<-rbind(firsty,secondy)

## question one
## delays in years
```{r}
### remove negative delay
firsty[firsty<0]<-NA
secondy[secondy<0]<-NA

### calculate monthly delay for ArrDelay and DepDelay
ArrDelay_1m<-aggregate(firsty$ArrDelay,list(firsty$Month),mean,na.rm=TRUE)
DepDelay_1m<-aggregate(firsty$DepDelay,list(firsty$Month),mean,na.rm=TRUE)
ArrDelay_2m<-aggregate(secondy$ArrDelay,list(secondy$Month),mean,na.rm=TRUE)
DepDelay_2m<-aggregate(secondy$DepDelay,list(secondy$Month),mean,na.rm=TRUE)

### rename column
ArrDelay_1m<-ArrDelay_1m%>%rename(Month = Group.1,ArrDelay = x)
DepDelay_1m<-DepDelay_1m%>%rename(Month = Group.1,DepDelay = x)
ArrDelay_2m<-ArrDelay_2m%>%rename(Month = Group.1,ArrDelay = x)
DepDelay_2m<-DepDelay_2m%>%rename(Month = Group.1,DepDelay = x)

### combine ArrDelay and DepDelay
total_delay1m=ArrDelay_1m%>%full_join(DepDelay_1m,by="Month")
total_delay2m=ArrDelay_2m%>%full_join(DepDelay_2m,by="Month")

### add year column in total_delay
Year<- 2007
total_delay1m<- cbind(total_delay1m,Year)
Year<- 2008
total_delay2m<- cbind(total_delay2m,Year)

### combine total_delay12
total_delay12m<-rbind(total_delay1m,total_delay2m)

### find trend in 2007 and 2008
library(ggplot2) 
ggplot() + 
geom_line(data=total_delay12m,aes(x=Month, y=ArrDelay,colour="ArrDelay"),size=1)+
geom_point(data=total_delay12m,aes(x=Month,y=ArrDelay,colour="ArrDelay"),size=3)+ 
geom_line(data=total_delay12m,aes(x=Month, y=DepDelay,colour="DepDelay"),size=1)+
geom_point(data=total_delay12m,aes(x=Month,y=DepDelay,colour="DepDelay"),size=3)+
labs(title = "Delay in Months",x="Month",y="Delay")+
facet_grid(~ Year)
```

## delays in days of week 
```{r}
### remove negative delay
df12[df12<0]<-NA

### caculate average delay
ArrDelay_12w<-aggregate(df12$ArrDelay,list(df12$DayOfWeek),mean,na.rm=TRUE)
DepDelay_12w<-aggregate(df12$DepDelay,list(df12$DayOfWeek),mean,na.rm=TRUE)

### rename column
ArrDelay_12w<-ArrDelay_12w%>%rename(DayOfWeek = Group.1,ArrDelay = x)
DepDelay_12w<-DepDelay_12w%>%rename(DayOfWeek = Group.1,DepDelay = x)

### combine ArrDelay and DepDelay
week_delay<-merge(ArrDelay_12w,DepDelay_12w,by="DayOfWeek",all=FALSE)

### melt week_delay
library("reshape2")
week_delay<- melt(week_delay,id.var = 'DayOfWeek')

### plot delay
library(ggplot2)
ggplot(week_delay, aes(fill=variable, y=value, x=DayOfWeek)) + 
    geom_bar(position="dodge", stat="identity")
```

## delay in days
```{r}
### remove negative delay
df12[df12<0]<-NA

### subset df_day
df_day<-df12[,c('DepTime','DepDelay','ArrTime','ArrDelay')]

### decide time interval
df_day$DepTime<- cut(df_day$DepTime,breaks=c(0,200,400,600,800,1000,1200,1400,1600,1800,2000,2200,2400),labels=c("2","4","6","8","10","12","14","16","18","20","22","24"),include.lowest=TRUE,right=FALSE)
df_day$ArrTime<- cut(df_day$ArrTime,breaks=c(0,200,400,600,800,1000,1200,1400,1600,1800,2000,2200,2400),labels=c("2","4","6","8","10","12","14","16","18","20","22","24"),include.lowest=TRUE,right=FALSE)

### calculate average delay
avg_depdelay<-df_day%>%
group_by(DepTime)%>%
summarise_at(vars(DepDelay),list(mean),na.rm=TRUE)
avg_arrdelay<-df_day%>%
group_by(ArrTime)%>%
summarise_at(vars(ArrDelay),list(mean),na.rm=TRUE)

### merge avg_delay
avg_delay<- merge(avg_depdelay,avg_arrdelay,by.x = "DepTime",by.y = "ArrTime")
avg_delay<- avg_delay[c(1,2,3,4,5,7,8,9,11,12),]

### plot line graph
library(ggplot2) 
ggplot() + 
geom_line(data=avg_delay,aes(x=DepTime, y=ArrDelay,colour="ArrDelay"),size=1)+
geom_point(data=avg_delay,aes(x=DepTime,y=ArrDelay,colour="ArrDelay"),size=3)+ 
geom_line(data=avg_delay,aes(x=DepTime, y=DepDelay,colour="DepDelay"),size=1)+
geom_point(data=avg_delay,aes(x=DepTime,y=DepDelay,colour="DepDelay"),size=3)+
labs(title = "Delay in Day",x="Hours",y="Delay")
```

## question two
## use graph 
```{r}
### prepare plane df
plane<-plane[,c(1,4)]
plane<-plane%>%
rename(TailNum = tailnum)

### prepare delay df
df12<-df12%>%filter(df12$Diverted!=0)
df12$total_delay<-apply(df12[,c(15,16)],1,sum,na.rm=TRUE)
delay<-df12[,c(11,30)]%>%
group_by(TailNum)%>%
summarise_at(vars(total_delay),list(mean),rm.na=TRUE)

### merge by tail number
plane_delay<-merge(plane,delay,by="TailNum")

### transfer datatype to date
plane_delay$issue_date<-as.Date(plane_delay$issue_date,'%m/%d/%Y')
summary(plane_delay)

### scatter plot
library(ggplot2)
ggplot(plane_delay,aes(x=issue_date,y=total_delay))+
geom_point()
```

## use two samples t-test 
## test variance with f-test
```{r}
### mutate new classification column 
plane_delay <- plane_delay %>% mutate(f_test= if_else(issue_date<'2001-10-09', "old", "new"))

### remove na
library(tidyr)
plane_delay<- plane_delay%>%drop_na()

### split into new and old df and make vectors
new <- plane_delay[plane_delay$f_test == 'new', ]
old <- plane_delay[plane_delay$f_test == 'old', ]
vector_new <- new$total_delay
vector_old <- old$total_delay

var.test(vector_new,vector_old)
```

## t-test
```{r}
t.test(vector_old,vector_new,alternative="greater",var.equal=TRUE)
```

## question three
## flights change over time
### assign two airports
```{r}
A<-function(airport){tryCatch(
  filter(df12,Origin==airport)
  ,error = function(e)
 print("please type correct airport code"))
}

B<-function(airport){tryCatch(
  filter(df12,Dest==airport)
  ,error = function(e)
 print("please type correct airport code"))
}
df12<-A("ATL")
df12<-B("DFW")
```

### months
```{r}
### df of monthly flights
month<-df12%>%count(Year,Month)

### plot chart
library(ggplot2) 
ggplot() + 
geom_line(data=month,aes(x=Month, y=n),size=1)+
geom_point(data=month,aes(x=Month,y=n),size=3)+ 
labs(title = "Flights in Months",x="Month",y="Number of Flights")+
facet_grid(~ Year)
```

### week
```{r}
### df of flights in week
week<-df12%>%count(DayOfWeek)

### plot chart
ggplot() + 
geom_line(data=week,aes(x=DayOfWeek, y=n),size=1)+
geom_point(data=week,aes(x=DayOfWeek,y=n),size=3)+ 
labs(title = "Flights in Day of Week",x="Day of Week",y="Number of Flights")
```

### day
```{r}
### decide time interval
df12$Time<- df12$CRSDepTime/100
df12$Time<-floor(df12$Time)

### df of flights in week
day<-df12%>%count(Time)

### plot chart
ggplot() + 
geom_line(data=day,aes(x=Time, y=n),size=1)+
geom_point(data=day,aes(x=Time,y=n),size=3)+ 
labs(title = "Flights in Day",x="Hours",y="Number of Flights")
```

## people movement between airports
### assign time period to df
```{r}
### assign year
year<-2007
if(year==2007 |year==2008){
  year<-year}else{print("please assign year from 2007 to 2008")
    stop()}

### assign month
month<-9
if(month>=1 & month<=12){
  month<-month}else{print("please assign month from 1 to 12")
    stop()}

### assign day
day<-0
if(day>=0 &day<=31){
  day<-day}else{print("please assign day from 0 to 31")
    stop()}
```

### prepare  df 
```{r}
### select data from assigned time frame 
df12<-rbind(firsty,secondy)
assign_time<-df12%>%filter(Year==year)%>%filter(Month==month)%>%filter(if(day%in%DayofMonth)DayofMonth==day else DayofMonth>0)

### create flight and count column
flights<-assign_time%>%mutate(flight=paste(Origin,Dest))%>%group_by(flight)%>%mutate(count=n())

### remove duplicate flight and sort flight
flights<-flights[!duplicated(flights$flight), ]

### extract important flights
flights<-flights%>%arrange(desc(count))
flights<-flights[1:20,]

### merge assign_time with airport_code
flights<-flights%>%select(Origin,Dest)%>% merge( airport_code, by.x = "Origin", by.y = "iata")%>%merge(airport_code, by.x = "Dest", by.y = "iata")%>% select(Origin,Dest,lat.x,long.x,lat.y,long.y)
```

### plot map
```{r}
library(ggplot2)
library(ggrepel)
usMap <- borders("state", colour="grey", fill="white")
title<-paste("Flights on ",month,"/",day,",",year)
allUSA <- ggplot() + usMap +
  geom_curve(data=flights,
             aes(x=long.x, y=lat.x, xend=long.y, yend=lat.y),
             col="#b29e7d",
             size=.7,
             curvature=0.2) +
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks=element_blank(),
        plot.title=element_text(hjust=0.5, size=12)) +
  ggtitle(title)

allUSA
```

## question four
### copy df12
dforigin<-rbind(firsty,secondy)

### assign airport A with ""
```{r}
A<-function(airport){tryCatch(
  filter(df12,Origin==airport)
  ,error = function(e)
 print("please type correct airport code"))
}
df12<-A("ATL")
```

## type one
```{r}
### depdelay in A and arrdelay in Bs from A
t1<-df12%>%group_by(Dest)%>%summarise(depdelay=mean(DepDelay,na.rm=TRUE),arrdelay_in_B=mean(ArrDelay,na.rm=TRUE))

### test correlation of depdelay and arrdelay_in_B in t1
cor(t1$depdelay, t1$arrdelay_in_B, method = "pearson")
cor.test(t1$depdelay, t1$arrdelay_in_B, method = "pearson")
```

## type two
```{r}
### depdelay in A to Bs 
t2<-df12%>%group_by(Dest)%>%summarize(depdelay=mean(DepDelay,na.rm=TRUE))

### depdelay in Bs to Cs
get_depdelay<-function(x){
 predf<-filter(dforigin,Origin==x)
  mean(predf$DepDelay,na.rm = TRUE)}

depdelay_in_B<-sapply(X=t2$Dest,FUN = get_depdelay)
t2$depdelay_in_B<-depdelay_in_B

### test correlation of depdelay and arrdelay_in_B in t1
cor(t2$depdelay, t2$depdelay_in_B, method = "pearson")
cor.test(t2$depdelay, t2$depdelay_in_B, method = "pearson")
```

## type three
```{r}
### arrdelay in Bs from A
t3<-df12%>%group_by(Dest)%>%summarize(arrdelay_in_B=mean(ArrDelay,na.rm=TRUE))

### arrdelay in Cs from Bs
get_arrdelay<-function(x){
 predf<-filter(dforigin,Origin==x)
  mean(predf$ArrDelay,na.rm = TRUE)}

arrdelay_in_C<-sapply(X=t3$Dest,FUN = get_arrdelay)
t3$arrdelay_in_C<-arrdelay_in_C

### test correlation of arrdelay_in_B and arrdelay_in_C in t3
cor(t3$arrdelay_in_B, t3$arrdelay_in_C, method = "pearson")
cor.test(t3$arrdelay_in_B, t3$arrdelay_in_C, method = "pearson")
```

## type four
```{r}
### arrdelay in Bs from A
t4<-df12%>%group_by(Dest)%>%summarize(arrdelay_in_B=mean(ArrDelay,na.rm=TRUE))

### depdelay in Cs to Ds
depdelay_in_C<-sapply(X=t4$Dest,FUN = get_depdelay)
t4$depdelay_in_C<-depdelay_in_C

### test correlation of arrdelay_in_B and depdelay_in_C in t4
cor(t4$arrdelay_in_B, t4$depdelay_in_C, method = "pearson")
cor.test(t4$arrdelay_in_B, t4$depdelay_in_C, method = "pearson")
```

## question five
## prepare data
```{r}
### dummy for airport
x<-df12$Origin%>%unique()
df12$Origin = factor(df12$Origin,levels =x) 
y<-df12$Dest%>%unique()
df12$Dest = factor(df12$Dest,levels =y) 

### dummy for carrier
z<-df12$UniqueCarrier%>%unique()
df12$UniqueCarrier = factor(df12$UniqueCarrier,levels =z)

### remove irrelevant column and na
library(tidyr)
df12<-df12%>%select(-TailNum,-Diverted,-Cancelled,-CancellationCode,-TaxiIn,-TaxiOut)
df12<-drop_na(df12)

### df for depdelay
df_D<-df12%>%select(-ArrDelay)

### df for arrdelay
df_A<-df12%>%select(-DepDelay)
```

## depdelay
## traing and test model
```{r}
### sampling
df_D<-df_D[sample(nrow(df_D),10000),]
library(caTools)
set.seed(123)
split = sample.split(df_D$DepDelay, SplitRatio = 0.8)

training_set = subset(df_D, split == TRUE)
test_set = subset(df_D, split == FALSE)

regressor_D = lm(formula = DepDelay ~ .,data = training_set)
summary(regressor_D)

### Predicting the test set results
test_set<-test_set%>%filter(Origin%in%training_set$Origin)%>%filter(Dest%in%training_set$Dest)
y_pred = predict(regressor_D, newdata = test_set)
```

## model accuracy 
```{r}
### plot pred_train and depdelay
training_set<-training_set%>%mutate(pred_train=predict(regressor_D))
plot(training_set$DepDelay,training_set$pred_train)
plot(test_set$DepDelay,y_pred)

### the accuracy of training_set (RMSE)
training_set<-training_set%>%mutate(error=pred_train-DepDelay,sq.error=error^2)
mse_train<-mean(training_set$sq.error)%>%sqrt() #6.99967101

### the accuracy of test_set (RMSE)
mse_test<-sqrt(mean((y_pred-test_set$DepDelay)^2))     #7.295295
```

## arrdelay
## traing and test model
```{r}
df_A<-df_A[sample(nrow(df_A),10000),]
library(caTools)
set.seed(123)
split = sample.split(df_A$ArrDelay, SplitRatio = 0.8)

training_set = subset(df_A, split == TRUE)
test_set = subset(df_A, split == FALSE)

regressor_A = lm(formula = ArrDelay ~ .,data = training_set)
summary(regressor_A) 

### Predicting the test set results
test_set<-test_set%>%filter(Origin%in%training_set$Origin)%>%filter(Dest%in%training_set$Dest)
y_pred = predict(regressor_A, newdata = test_set)
```

## model accuracy 
```{r}
### plot pred_train and depdelay
training_set<-training_set%>%mutate(pred_train=predict(regressor_A))
plot(training_set$ArrDelay,training_set$pred_train)
plot(test_set$ArrDelay,y_pred)

### the accuracy of training_set (RMSE)
training_set<-training_set%>%mutate(error=pred_train-ArrDelay,sq.error=error^2)
mse_train<-mean(training_set$sq.error)%>%sqrt() #6.971924564

### the accuracy of test_set (RMSE)
mse_test<-sqrt(mean((y_pred-test_set$ArrDelay)^2))     #7.465989403
```



